---
title: "ST558- Project 3"
author: "Rohan Prabhune, Manan Shah"
output: 
  github_document:
    toc: true
    toc_depth: 2
always_allow_html: true
---
```{r Temp1,echo=FALSE,eval=TRUE,message=FALSE,warning=FALSE}
knitr::opts_chunk$set(fig.align = 'center',echo=TRUE,eval=TRUE,message=FALSE,warning=FALSE) 
```

# Loading the necessary packages
```{r packages}
library(tidyverse)
library(caret)
```

# Reading the data:
*Citation for the dataset:  
K. Fernandes, P. Vinagre and P. Cortez. A Proactive Intelligent Decision
Support System for Predicting the Popularity of Online News. Proceedings
of the 17th EPIA 2015 - Portuguese Conference on Artificial Intelligence,
September, Coimbra, Portugal.*

```{r Data}
#Reading the data
dataset <- read_csv("data/OnlineNewsPopularity.csv",show_col_types=FALSE) 
```

# Pre-processing:  

## Subset data based on type of article
```{r preprocessing}
# Vector of columns other than data_channel_is*
remaining_cols = c(names(dataset))[c(1:13,20:61)]
# Subset based on type of article
input_data_channel = "entertainment"
select_var = paste0("data_channel_is_",input_data_channel)
subset_cols = append(remaining_cols,select_var)
# Subset data
df <- dataset %>% select(all_of(subset_cols)) %>%
  filter(dataset[select_var] == 1) %>%
  select(all_of(remaining_cols))
```

## Variable selection
```{r preprocessing2}
#Remove url and timedelta
# Remove "weekday_is_saturday" and "weekday_is_sunday" to only keep "is_weekend"
df <- df %>% select(-url,-timedelta,-weekday_is_saturday,-weekday_is_sunday)
df
```