---
title: "Automate R markdown"
author: "Rohan Prabhune, Manan Shah"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE,eval=TRUE,message=FALSE,warning=FALSE)
```

# Automating R markdown  
`out_params` is a vector of parameters which are used as input parameters to **main.Rmd** file. Since the `param` argument in `render()` function requires params to be a named list, we have converted each element of `out_params` to a named list using `lapply()` into `out_params_list`.  

`out_filename` is a vector of output file names we want for each of the input parameter.  

Since there are only six parameters, we decided to go with a for loop to iterate over the file names and vector of named lists to create six different .md files based on the type of article.  
```{r automate}
library(rmarkdown)

out_params <- c("lifestyle","entertainment","bus","socmed","tech","world")
out_params_list = lapply(out_params, FUN = function(x){list(var = x)})

out_filename <- c("Lifestyle_analysis.md","Entertainment_analysis.md",
                  "Business_analysis.md","Social_media_analysis.md",
                  "Tech_analysis.md","World_analysis.md")
for (i in 1:6){
  rmarkdown::render(input = "main.Rmd", output_file = out_filename[i],
                    params = out_params_list[[i]])
}
```
